A Large-Scale Study of Polymorphism in R

### 2019-04-06 paper/main.tex
related

\begin{figure}[!hb]{\small\begin{lstlisting}[style=R]
fixGenes <- function (network, fixIndices, values) {
  ...
  network$fixed[fixIndices] <- as.integer(values)
  ...
\end{lstlisting}}\caption{Example of \C/\D argument (from {\tt BoolNet} package).}\label{fig:chardbl}\end{figure}

@@

The idea of measuring polymorphism via some program analysis isn't new.
For instance, some work~\cite{aakerblom2015measuring} measures the polymorphism of Python programs, but with a slightly different goal than ours: authors seek to assess the need for expressiveness in type systems that are retrofit onto dynamic languages.
Interestingly, they draw some similar conclusions to us: they find that monomorphism is far more frequent than polymorphism, and even find that typically very few functions are very highly polymorphic.

More generally, there is an existing literature on analyzing usage patterns of language features.  
For instance, the dynamic features of Smalltalk were analyzed in some work~\cite{callau2011howdevelopers}, which aimed to see how often the highly dynamic capabilities of Smalltalk were used and in which kinds of projects.

Other analysis work~\cite{milojkovic2017duck} ... .

The type system we propose in this work could be incorporated into R as a {\it gradual type system}.
Gradual type systems~\cite{SiekTaha06} aim to bridge the gap between dynamic and static typing by allowing programmers to gradually add type annotations to their programs.
Gradually-typed languages are becoming increasingly popular, with languages such as TypeScript~\cite{typescript13} becoming more mainstream.
...

>>>>

\begin{figure}[!hb]{\small\begin{lstlisting}[style=R]
fixGenes <- function (network, fixIndices, values) {
  ...
  network$fixed[fixIndices] <- as.integer(values)
  ...
\end{lstlisting}}\caption{Example of \C/\D argument (from {\tt BoolNet} package).}\label{fig:chardbl}\end{figure}%$

@@

The idea of measuring polymorphism via some program analysis is not new.
For instance, \citep{aakerblom2015measuring} measured the polymorphism of
Python programs. Their goals were related to our, they wanted to assess the
need for expressiveness in type systems that are retrofitted onto dynamic
languages.  Interestingly, they drew similar conclusions: monomorphism is
far more frequent than polymorphism, and very few functions are very highly
polymorphic.  More generally, the dynamic features of Smalltalk were
analyzed by~\citep{callau2011howdevelopers} and duck typing was studied
by~\citep{milojkovic2017duck}.  Adding type systems to untyped language was
proposed by~\citep{SiekTaha06} who aim to bridge the gap between dynamic and
static typing.  Gradually-typed languages such as
TypeScript~\cite{typescript13} are pointing the way to what we could do
for R.
